Generating Correlated Data for Omics Simulation

### 2025-06-18 manuscript/paper.qmd
Revise CYCLOPS comment

By default, CYCLOPS performs dimension reduction so that each dimension (called an "eigengene") contains at least 3% of the total variance.
We found that CYCLOPS accuracy depended significantly on this parameter, with the default producing good accuracy across all simulations.
However, when dropping CYCLOPS to require just 2% variance in each eigengene, we found that its accuracy depends significantly on the dependence structure of the simulated time series data (@fig-CYCLOPS a).
At that setting, CYCLOPS accuracy is much higher in the PCA method and moderately improved in Wishart method, compared to the 'independent' method. 
This difference is likely driven by the difference in the number of eigengenes used (@fig-CYCLOPS b), which is a measure of how much dependence is present in the dataset.
This demonstrates that the correlation structure of the transcriptome can have a major impact on accuracy.
In this application, the differences between corpcor and Wishart were such that, while Wishart improved upon the number of eigengenes used compared to corpcor, both undershot the level of correlation in the real data set (see Supplemental Figure 2 d) and the difference between them was too small to affect the actual accuracy of CYCLOPS as measured by the circular correlation.

@@

```{r}
#| label: fig-CYCLOPS
#| fig-cap: "Performance of CYCLOPS on simulated time series datasets based on mouse cortex dataset (GSE151565), using eigengenes of at least 2% variance. (a) Absolute circular correlations between true phases and CYCLOPS estimated phases on the simulated datasets. (b) The number of eigengenes used by CYCLOPS; the dotted line indicates the number of eigengenes used by CYCLOPS on the real data (5). (c-f) Examples of CYCLOPS estimated phases on the simulated datasets. CYCLOPS shows good performance when it separates out points by color (true circadian time)."
#| fig-width: 12
#| fig-height: 8
cyclops_plot
```

>>>>

By default, CYCLOPS performs dimension reduction so that each dimension (called an "eigengene") contains at least 3% of the total variance.
We found that CYCLOPS accuracy depended significantly on this parameter, with the default producing good accuracy across all simulations.
However, when dropping CYCLOPS to require just 2% variance in each eigengene, we found that its accuracy depends significantly on the dependence structure of the simulated time series data (@fig-CYCLOPS a).
At that setting, CYCLOPS accuracy is much higher in the PCA method and moderately improved in Wishart method, compared to the 'independent' method. 
This difference is likely driven by the difference in the number of eigengenes used (@fig-CYCLOPS b), which is a measure of how much dependence is present in the dataset.
Notably, the number of eigengenes and accuracy tracked both with each other and with the amount of gene-gene correlation (see Supplemental Figure 3 d) across the methods.
This demonstrates that the correlation structure of the transcriptome can have a major impact on accuracy.

@@

```{r}
#| label: fig-CYCLOPS
#| fig-cap: "Accuracy of CYCLOPS on simulated time series datasets based on mouse cortex dataset (GSE151565), using eigengenes of at least 2% variance. (a) Absolute circular correlations between true phases and CYCLOPS estimated phases on the simulated datasets. (b) The number of eigengenes used by CYCLOPS; the dotted line indicates the number of eigengenes used by CYCLOPS on the real data (5). (c-f) Examples of CYCLOPS estimated phases on the simulated datasets. CYCLOPS shows good performance when it separates out points by color (true circadian time)."
#| fig-width: 12
#| fig-height: 8
cyclops_plot
```
